---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
```{r setup, include = FALSE}
require(qfratio)
set.seed(64501)
```

# qfratio: R Package for Moment of Ratios of Quadratic Forms

<!-- badges: start -->
<!-- badges: end -->


This package provides functions to evaluate moments of ratios of quadratic forms in normal variables, specifically using recursive algorithms developed by Chikuse, Hillier, Bao, Kan and colleagues.

There exists a couple of `Matlab` programs developed by Raymond Kan (available from <https://www-2.rotman.utoronto.ca/~kan/>), but this `R` package has been developed independently with added functionalities. This has originally been developed for a biological application, specifically for evaluating average evolvability measures in quantitative genetics.

This project is under active development. Most planned functionalities have been implemented, but have not been extensively documented. At present, substantial restructuring might be possible.



## Installation

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("watanabe-j/qfratio")
```

This package has the following dependencies:

    Imports: Rcpp (>= 1.0.8.3)
    LinkingTo: Rcpp, RcppEigen
    Suggests: gsl, mvtnorm


## Examples

Here are some simple examples:

```{r}
## Simple matrices
nv <- 4
A <- diag(1:nv)
B <- diag(sqrt(nv:1))

## Expectation of (x^T A x)^2 / (x^T x)^2 where x ~ N(0, I)
qfrm(A, p = 2)

## Compare with Monte Carlo mean
mean(rqfr(1000, A = A, p = 2))

## Expectation of (x^T A x)^1/2 / (x^T x)^1/2
(mom_A0.5 <- qfrm(A, p = 1/2))

## Monte Carlo mean
mean(rqfr(1000, A = A, p = 1/2))

plot(mom_A0.5)

## Expectation of (x^T x) / (x^T A^-1 x)
##   = "average conditional evolvability"
(avr_cevoA <- qfrm(diag(nv), solve(A)))

mean(rqfr(1000, A = diag(nv), B = solve(A), p = 1))
plot(avr_cevoA)

## Expectation of (x^T x)^2 / (x^T A x) (x^T A^-1 x) 
##   = "average autonomy"
(avr_autoA <- qfmrm(diag(nv), A, solve(A), p = 2, q = 1, r = 1))

mean(rqfmr(1000, A = diag(nv), B = A, D = solve(A), p = 2, q = 1, r = 1))
plot(avr_autoA)

## Expectation of (x^T A B x) / ((x^T A^2 x) (x^T B^2 x))^1/2 
##   = "average vector correlation"
## whose Monte Carlo evaluation is called the "random skewers" analysis,
## while this is essentially an analytic solution (with slight truncation error)
(avr_rcorA <- qfmrm(crossprod(A, B), crossprod(A), crossprod(B), 
                    p = 1, q = 1/2, r = 1/2))

mean(rqfmr(1000, A = crossprod(A, B), B = crossprod(A), D = crossprod(B), 
           p = 1, q = 1/2, r = 1/2))
plot(avr_rcorA)


## More complex (but arbitrary) example
## Expectation of (x^T A x)^2 / (x^T B x)^3 where x ~ N(mu, Sigma)
mu <- 1:nv / nv
Sigma <- diag(runif(nv) * 3)
(mom_A2B3 <- qfrm(A, B, p = 2, q = 3, mu = mu, Sigma = Sigma, 
                  m = 500, use_cpp = TRUE))
plot(mom_A2B3)
```


## References

Bao, Y. & Kan, R. (2013). On the moments of ratios of quadratic forms in normal random variables. *Journal of Multivariate Analysis*, **117**, 229--245. doi: [10.1016/j.jmva.2013.03.002](https://doi.org/10.1016/j.jmva.2013.03.002).

Hillier, G., Kan, R, & Wang, X. (2009). Computationally efficient recursions for top-order invariant polynomials with applications. *Econometric Theory*, **25**, 211--242. doi: [10.1017/S0266466608090075](https://doi.org/10.1017/S0266466608090075).

Hillier, G., Kan, R, & Wang, X. (2014). Generating functions and short recursions, with applications to the moments of quadratic forms in doi: [10.1017/S0266466613000364](https://doi.org/10.1017/S0266466613000364).

Smith, M. D. (1989). On the expectation of a ratio of quadratic forms in normal variables. *Journal of Multivariate Analysis*, **31**, 244--257. doi: [10.1016/0047-259X(89)90065-1](https://doi.org/10.1016/0047-259X(89)90065-1)

Smith, M. D. (1993). Expectations of ratios of quadratic forms in normal variables: evaluating some top-order invariant polynomials. *Australian Journal of Statistics*, **35**, 271--282. doi: [10.1111/j.1467-842X.1993.tb01335.x](https://doi.org/10.1111/j.1467-842X.1993.tb01335.x).
