---
output: github_document
bibliography: vignettes/bibliography.bib
link-citations: TRUE
nocite: |
  @BaoKan2013
  @HillierEtAl2009
  @HillierEtAl2014
  @Smith1989
  @Smith1993
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
```{r setup, include = FALSE}
require(qfratio)
set.seed(64501)
```

# qfratio: R Package for Moments of Ratios of Quadratic Forms

<!-- badges: start -->
<!-- badges: end -->


This package provides functions to evaluate moments of ratios
(and products) of quadratic forms in normal variables, specifically using
recursive algorithms developed by Hillier, Bao, Kan and colleagues.
Generating functions for these moments are closely related to the
top-order zonal and invariant polynomials of matrix arguments.

There exist a couple of `Matlab` programs developed by Raymond Kan
(available from <https://www-2.rotman.utoronto.ca/~kan/>), but this
`R` package is an independent project (not a fork or translation)
and has different functionalities, including evaluation of moments of
multiple ratios of a particular form and scaling to avoid numerical overflow.
This has originally been developed for a biological application,
specifically for evaluating average evolvability measures in
evolutionary quantitative genetics [@Watanabe2022cevo].

This project is in a late stage of development.
Most functionalities have been implemented and are considered stable.
Documentations of the package and functions are partly under construction.
Feedbacks are much appreciated.



## Installation

***WARNING*** Installation size of this package can be very large
(~130 MB on Linux; ~3 MB on Windows with `Rtools42`),
as it involves lots of `RcppEigen` functions.

```{r eval = FALSE}
## Install devtools first:
# install.packages("devtools")

## Recommended installation (pandoc required):
devtools::install_github("watanabe-j/qfratio", dependencies = TRUE, build_vignettes = TRUE)

## Minimal installation:
# devtools::install_github("watanabe-j/qfratio")
```

This package has the following dependencies:

    Imports: Rcpp, MASS, rlang (>= 0.4.7)
    LinkingTo: Rcpp, RcppEigen
    Suggests: gsl, mvtnorm, graphics, stats, testthat (>= 3.0.0), knitr, rmarkdown

You also need [`pandoc`](https://pandoc.org) for correctly building
the vignette. (For `pandoc < 2.11`, `pandoc-citeproc` is required as well.)

## Examples

Here are some simple examples:

```{r}
## Simple matrices
nv <- 4
A <- diag(1:nv)
B <- diag(sqrt(nv:1))

## Expectation of (x^T A x)^2 / (x^T x)^2 where x ~ N(0, I)
qfrm(A, p = 2)

## Compare with Monte Carlo mean
mean(rqfr(1000, A = A, p = 2))

## Expectation of (x^T A x)^1/2 / (x^T x)^1/2
(mom_A0.5 <- qfrm(A, p = 1/2))

## Monte Carlo mean
mean(rqfr(1000, A = A, p = 1/2))

plot(mom_A0.5)

## Expectation of (x^T x) / (x^T A^-1 x)
##   = "average conditional evolvability"
(avr_cevoA <- qfrm(diag(nv), solve(A)))

mean(rqfr(1000, A = diag(nv), B = solve(A), p = 1))
plot(avr_cevoA)

## Expectation of (x^T x)^2 / (x^T A x) (x^T A^-1 x)
##   = "average autonomy"
(avr_autoA <- qfmrm(diag(nv), A, solve(A), p = 2, q = 1, r = 1))

mean(rqfmr(1000, A = diag(nv), B = A, D = solve(A), p = 2, q = 1, r = 1))
plot(avr_autoA)

## Expectation of (x^T A B x) / ((x^T A^2 x) (x^T B^2 x))^1/2
##   = "average response correlation"
## whose Monte Carlo evaluation is called the "random skewers" analysis,
## while this is essentially an analytic solution (with slight truncation error)
(avr_rcorA <- qfmrm(crossprod(A, B), crossprod(A), crossprod(B),
                    p = 1, q = 1/2, r = 1/2))

mean(rqfmr(1000, A = crossprod(A, B), B = crossprod(A), D = crossprod(B),
           p = 1, q = 1/2, r = 1/2))
plot(avr_rcorA)


## More complex (but arbitrary) example
## Expectation of (x^T A x)^2 / (x^T B x)^3 where x ~ N(mu, Sigma)
mu <- 1:nv / nv
Sigma <- diag(runif(nv) * 3)
(mom_A2B3 <- qfrm(A, B, p = 2, q = 3, mu = mu, Sigma = Sigma,
                  m = 500, use_cpp = TRUE))
plot(mom_A2B3)
```


## References
