---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
```{r setup, include = FALSE}
require(qfratio)
set.seed(64501)
```

# qfratio: R Package for Moment of Ratios of Quadratic Forms

<!-- badges: start -->
<!-- badges: end -->


This package provides functions to evaluate moments of ratios
(and products) of quadratic forms in normal variables, specifically using
recursive algorithms developed by Hillier, Bao, Kan and colleagues.
Generating functions for these moments are closely related to the
top-order zonal and invariant polynomials of matrix arguments.

There exist a couple of `Matlab` programs developed by Raymond Kan
(available from <https://www-2.rotman.utoronto.ca/~kan/>), but this
`R` package has been developed independently with different
functionalities.
This has originally been developed for a biological application,
specifically for evaluating average evolvability measures in
evolutionary quantitative genetics (Watanabe, 2022).

This project is **under active development**.
Most planned functionalities have been implemented, but some may lack
comprehensive documentation.
Substantial restructuring may happen in the future.



## Installation

***WARNING*** Installation size of this package can be very large
(~100 MB on Linux; ~3 MB on Windows with `Rtools42`),
as it involves lots of `RcppEigen` functions.

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("watanabe-j/qfratio")
```

This package has the following dependencies:

    Imports: Rcpp, MASS, rlang (>= 0.4.7)
    LinkingTo: Rcpp, RcppEigen
    Suggests: gsl, mvtnorm, graphics, stats, testthat (>= 3.0.0)

It is recommended to have the packages `gsl` and `mvtnorm` installed
for full functionality.
You can use `install_github(..., dependencies = TRUE)` to automatically
install the suggested packages.

## Examples

Here are some simple examples:

```{r}
## Simple matrices
nv <- 4
A <- diag(1:nv)
B <- diag(sqrt(nv:1))

## Expectation of (x^T A x)^2 / (x^T x)^2 where x ~ N(0, I)
qfrm(A, p = 2)

## Compare with Monte Carlo mean
mean(rqfr(1000, A = A, p = 2))

## Expectation of (x^T A x)^1/2 / (x^T x)^1/2
(mom_A0.5 <- qfrm(A, p = 1/2))

## Monte Carlo mean
mean(rqfr(1000, A = A, p = 1/2))

plot(mom_A0.5)

## Expectation of (x^T x) / (x^T A^-1 x)
##   = "average conditional evolvability"
(avr_cevoA <- qfrm(diag(nv), solve(A)))

mean(rqfr(1000, A = diag(nv), B = solve(A), p = 1))
plot(avr_cevoA)

## Expectation of (x^T x)^2 / (x^T A x) (x^T A^-1 x)
##   = "average autonomy"
(avr_autoA <- qfmrm(diag(nv), A, solve(A), p = 2, q = 1, r = 1))

mean(rqfmr(1000, A = diag(nv), B = A, D = solve(A), p = 2, q = 1, r = 1))
plot(avr_autoA)

## Expectation of (x^T A B x) / ((x^T A^2 x) (x^T B^2 x))^1/2
##   = "average vector correlation"
## whose Monte Carlo evaluation is called the "random skewers" analysis,
## while this is essentially an analytic solution (with slight truncation error)
(avr_rcorA <- qfmrm(crossprod(A, B), crossprod(A), crossprod(B),
                    p = 1, q = 1/2, r = 1/2))

mean(rqfmr(1000, A = crossprod(A, B), B = crossprod(A), D = crossprod(B),
           p = 1, q = 1/2, r = 1/2))
plot(avr_rcorA)


## More complex (but arbitrary) example
## Expectation of (x^T A x)^2 / (x^T B x)^3 where x ~ N(mu, Sigma)
mu <- 1:nv / nv
Sigma <- diag(runif(nv) * 3)
(mom_A2B3 <- qfrm(A, B, p = 2, q = 3, mu = mu, Sigma = Sigma,
                  m = 500, use_cpp = TRUE))
plot(mom_A2B3)
```


## References

Bao, Y. & Kan, R. (2013). On the moments of ratios of quadratic forms in
normal random variables. *Journal of Multivariate Analysis*, **117**,
229--245.
doi: [10.1016/j.jmva.2013.03.002](https://doi.org/10.1016/j.jmva.2013.03.002).

Hillier, G., Kan, R, & Wang, X. (2009). Computationally efficient
recursions for top-order invariant polynomials with applications.
*Econometric Theory*, **25**, 211--242.
doi: [10.1017/S0266466608090075](https://doi.org/10.1017/S0266466608090075).

Hillier, G., Kan, R, & Wang, X. (2014). Generating functions and short
recursions, with applications to the moments of quadratic forms in
noncentral normal vectors. *Econometric Theory*, **30**, 436--473.
doi: [10.1017/S0266466613000364](https://doi.org/10.1017/S0266466613000364).

Smith, M. D. (1989). On the expectation of a ratio of quadratic forms in
normal variables. *Journal of Multivariate Analysis*, **31**, 244--257.
doi: [10.1016/0047-259X(89)90065-1](https://doi.org/10.1016/0047-259X(89)90065-1)

Smith, M. D. (1993). Expectations of ratios of quadratic forms in normal
variables: evaluating some top-order invariant polynomials.
*Australian Journal of Statistics*, **35**, 271--282.
doi: [10.1111/j.1467-842X.1993.tb01335.x](https://doi.org/10.1111/j.1467-842X.1993.tb01335.x).

Watanabe, J. (2022). Exact expressions and numerical evaluation of average
evolvability measures for characterizing and comparing **G** matrices.
*bioRxiv* preprint, 2022.11.02.514929.
doi: [10.1101/2022.11.02.514929](https://doi.org/10.1101/2022.11.02.514929).
